---
title: "Bay Area"
subtitle: "🚴 <br/>Bike Share"
author: "Team 2: Five States"
date: "2017/10/01"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(servr.daemon = TRUE)
knitr::opts_chunk$set(echo = F, tidy = F, warning = F, message = F,
                      comment = "", tidy = F, fig.width = 10,
                      eval = require("DT", "tidyverse"))
library(tidyverse)
library(gridExtra)
library(DT)
```

# Background

- Bay Area Bike Share offers cheap, fast and on-demand service for the users in San Francisco who need commute to work or school, run errands and explore the city.  

- How to utilize data to optimize the operation which in turn to streamline the process and supply chain in order to satisfy customers’ demands by analyzing the data that the company has would be the top priority of the company.

- We have tracking the usage of the rented bike in the past years which makes it available to analyze by using the historical data.

.footnote[
  Team member: Peng Gong, Qiaozhi Jiang, Xinyu Mou, Zhanfeng Shen, Liangquan Zhou
]
???

here is some thing in the notes

---

# Objects and Approach

1. Exploratory analysis:
  
  - When do people use the bike?
  
  - How long do people use the bike?
  
  - Which stations has more usage of bike?
  
2. Linear Regression:
  How does the weather affect the bike usage?

---

# Data

Data comes from [Ford GoBike](http://www.bayareabikeshare.com/datachallenge) open data source.

Contains 4 tables:

  1. station.csv
  
  2. trip.csv
  
  3. weather.csv
  
  4. status.csv 

We use the data from 08/2015 - 08/2016

---

### station.csv

67 records – station ID, name, latitude, longitude, dockcount, city, installation date

```{r}
station <- read_csv("./../data/201608_station_data.csv") 
DT::datatable(
  head(station, 10),
  fillContainer = F, options = list(pageLength = 3, scrollX = T, dom = "tp"),
  rownames = F
)
```


---

### trip.csv

Approx. 314,000 records of individual trips

```{r}
trip <- read_csv("./../data/201608_trip_data.csv")
DT::datatable(
  head(trip, 10),
  fillContainer = F, options = list(pageLength = 3, scrollX = T, dom = "tp"),
  rownames = F
)
```

---

### weather.csv

1,830 records of daily weather by city

```{r}
weather <- read_csv("./../data/201608_weather_data.csv")
DT::datatable(
  head(weather, 10),
  fillContainer = F, options = list(pageLength = 3, scrollX = T, dom = "tp"),
  rownames = F
)
```

---
class: inverse, center, middle

# Exploratory Analysis

---

## Time and bike usage

```{r}
p1
```

???
with 1 year data - how does the bike usage affected the dates
 - weekday and weekend will affect
 - seems like season will affect, but actually is holiday (red dots with small value)

---

## Bike usage among a week

```{r}
p2
```

---

## Bike usage in a day

```{r}
p3
```

???

peak hours has the most usage

will weekday/weekends affect the usage in a day?

---

## Bike usage in a day

```{r}
p4
```

---

## Trip duration 

```{r}
gridExtra::grid.arrange(p5, p6, ncol = 1)
```

---

## Locations

```{r out.width='100%', fig.height=6, eval=require('leaflet')}
l8
```

---

## Bike usage in a city

Take San Jose as an example
```{r}
p7
```

---
class: inverse, center, middle

# Regression

## How does the weather affect the bike usage?

---

### weather.csv

```{r}
weather %>% glimpse
```

---

## Variable Selection

Response variable: `num of trips`

Possible predictor variables:

- `weekday`
- `mean_temperature_f`
- `precipitation_in`
- `mean_visibility_miles`
- `mean_humidity`
- `mean_wind_speed_mph`
- `cloud_cover`
- `events`

Use `regsubsets` function in `leaps` package for variable selection. Use San Francisco data.

---

## Diagnostic

```{r}
par(mfrow = c(2, 2), oma = c(0, 0, 2, 0))
plot(fit)
```

???

Residual Diagnostic:

1. Residual vs fitted has a pattern - transform on Y

1. qq plot - not normal - may be affected by extreme values

---

## Diagnostic

```{r}
p10
```

???

other factors could make model fail:

Seperated by weekday/weekend may not accurate. Should include national holidays 

---

## Diagnostic

Check weekdays with low usage
```{r}
DT::datatable(
  w1 %>% slice(1:18),
  fillContainer = F, options = list(pageLength = 5, scrollY = T, dom = "tp"),
  rownames = F
)
```

--

Treat national holidays as weekend, and we can get the model:

$$\sqrt{num\ of\ trips} = 18.685 - 35.562 weekend + 0.366 mean\ temperature +\\
1.534mean\ visibility\ miles - 5.91 events\ Rain - 5.767 events\ Rain\ Thunderstorm$$

???

we should exclude holidays. 

---

## Final Model

```{r}
summary(fit_holiday_adjust)
```

---

## Final Model 

```{r}
par(mfrow = c(2, 2), oma = c(0, 0, 2, 0))
plot(fit_holiday_adjust)
```

---
class: inverse, center, middle

# Thank You!
